Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion

### 2023-07-14 Manuscript.Rmd
references

---
title: "Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion"
author: 
  - name : "Alexandra G. Mitchell"
    affiliation : "1"
  - name : "Jesper Fischer Ehmsen"
    affiliation : "1"
  - name : "Daniel Elmstrom Christensen"
    affiliation : "2"
  - name : "Anna Villaume Stuckert"
    affiliation : "3"
  - name : "Patrick Haggard"
    affiliation : "4"  
  - name : "Francesca Fardo"
    affiliation : "1"

affiliation:
  - id : "1"
    institution : "Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University, Denmark"
  - id : "2"
    institution : "Interacting Minds Center (IMC), Department XXX, Aarhus University, Denmark"
  - id : "3"
    institution : "Departement of Neuroscience, University of Copenhagen, Denmark"
  - id : "4"
    institution : "Institute of Cognitive Neuroscience, University College London, United Kingdom"
  - id : "5"
    institution : "Danish Pain Research Center, Aarhus University, Denmark"
#date: "`r Sys.Date()`"
output:
  word_document:
  html_document: default
  pdf_document:
    latex_engine: xelatex
bibliography: knitting files/Citations.bib
link-citations: yes
linkcolor: blue
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: sentence
---

@@

The TGI is often described as encompassing two distinct perceptual components - an illusion of heat and an illusion of pain [@defrin_spatial_2008, @fardo_beyond_2020]. The illusion of heat, also known as synthetic heat [@fruhstorfer_significance_2003; @green_localization_1977; -@green_synthetic_2002; -@green_temperature_2004], refers to non-painful sensations evoked by the thermal grill [@defrin_spatial_2008; @kern_pharmacological_2008; -@kern_effects_2008; @bouhassira_investigation_2005; @adam_relationships_2014].
The illusion of pain, which is the most recognised aspect of the TGI, is the distinctive burning sensation that accompanies the simultaneous presentation of cold and warm stimuli [@craig_thermal_1994; -@craig_functional_1996; @bach_thermal_2011].
The hallmark of both illusory components is a qualitative change in perception when the cold and warm stimuli are applied concurrently, compared to when they are presented individually.
Historically, the thermosensory and painful components of the TGI were explained by distinct spinal and supraspinal mechanisms, respectively.
The enhanced perception of heat in TGI was explained by a spinal inhibitory mechanism, drawing from observations in an animal model where simultaneous application of cold and warm temperatures reduced the activity of cold-specific spinal neurons compared to when cold was applied alone [@craig_thermal_1994]. Instead, the illusory pain component was ascribed to a disinhibition mechanism at the level of the thalamus, primarily based on the observations of unremitting pain following thalamic lesions [@craig_thermal_1994; -@bud_craig_new_1998].

>>>>

---
title: "Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion"
author: 
  - name : "Alexandra G. Mitchell"
    affiliation : "1"
  - name : "Jesper Fischer Ehmsen"
    affiliation : "1"
  - name : "Daniel Elmstrom Christensen"
    affiliation : "2"
  - name : "Anna Villaume Stuckert"
    affiliation : "3"
  - name : "Patrick Haggard"
    affiliation : "4"  
  - name : "Francesca Fardo"
    affiliation : "1"  
affiliation:
  - id : "1"
    institution : "Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University, Denmark"
  - id : "2"
    institution : "Interacting Minds Center (IMC), Department XXX, Aarhus University, Denmark"
  - id : "3"
    institution : "Departement of Neuroscience, University of Copenhagen, Denmark"
  - id : "4"
    institution : "Institute of Cognitive Neuroscience, University College London, United Kingdom"
  - id : "5"
    institution : "Danish Pain Research Center, Aarhus University, Denmark"
#date: "`r Sys.Date()`"
output:
  word_document:
  html_document: default
  pdf_document:
    latex_engine: xelatex
bibliography: knitting files/Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion.bib
link-citations: yes
linkcolor: blue
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: sentence
---

@@

The TGI is often described as encompassing two distinct perceptual components - an illusion of heat and an illusion of pain [@defrin_spatial_2008, @fardo_beyond_2020]. The illusion of heat, also known as synthetic heat [@fruhstorfer_significance_2003; @green_localization_1977; -@green_synthetic_2002; -@green_temperature_2004], refers to non-painful sensations evoked by the thermal grill [@defrin_spatial_2008; @kern_pharmacological_2008; -@kern_effects_2008; @bouhassira_investigation_2005; @adam_relationships_2014].
The illusion of pain, which is the most recognised aspect of the TGI, is the distinctive burning sensation that accompanies the simultaneous presentation of cold and warm stimuli [@craig_thermal_1994; -@craig_functional_1996; @bach_thermal_2011].
The hallmark of both illusory components is a qualitative change in perception when the cold and warm stimuli are applied concurrently, compared to when they are presented individually.
Historically, the thermosensory and painful components of the TGI were explained by distinct spinal and supraspinal mechanisms, respectively.
The enhanced perception of heat in TGI was explained by a spinal inhibitory mechanism, drawing from observations in an animal model where simultaneous application of cold and warm temperatures reduced the activity of cold-specific spinal neurons compared to when cold was applied alone [@craig_thermal_1994]. Instead, the illusory pain component was ascribed to a disinhibition mechanism at the level of the thalamus, primarily based on the observations of unremitting pain following thalamic lesions [@craig_thermal_1994; -@craig_new_1998].
